---
title: "Practical Machine Learning - Prediction Assignment Writeup"
author: "David Hinostroza Vintes"
date: "17 de agosto de 2016"
output: html_document
---

For this assignment I analyzed the provided data to determine what activity an individual perform. To do this I made use of caret and randomForest, this allowed me to generate correct answers for each of the 20 test data cases provided in this assignment.

# Install packages and load the required libraries
```{r}
set.seed(1234)
library(lattice); 
library(ggplot2); 
library(caret); 
library(randomForest); 
library(rpart); 
library(rpart.plot);
library(rattle)
library(ROCR)
```

## The location where the training data is to be downloaded from
```{r}
trainUrl <-"http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
```
## The location where the testing data is to be downloaded from
```{r}
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

## data load and clean up
```{r}
trainingset <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testingset <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
```

## Perform exploratory analysis
```{r}
dim(trainingset); 
dim(testingset); 
summary(trainingset); 
summary(testingset); 
str(trainingset); 
str(testingset); 
head(trainingset); 
head(testingset)
```
## Take a look at the Training data classe variable
```{r}
summary(trainingset$classe)
```

## Delete columns with all missing values
```{r}
trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset <-testingset[,colSums(is.na(testingset)) == 0]
```

## Drop the first 7 variables because these are made up of metadata that would cause the model to perform poorly
```{r}
trainingset <-trainingset[,-c(1:7)]
testingset <-testingset[,-c(1:7)]
```
## Check NearZeroVariance variables
```{r}
nearZeroVar(trainingset, saveMetrics=TRUE)
nearZeroVar(testingset, saveMetrics=TRUE)
```

# Partitioning the data for Crossvalidation
## partition the data so that 75% of the training dataset into training and the remaining 25% to testing
```{r}
traintrainset <- createDataPartition(y=trainingset$classe, p=0.75, list=FALSE)
TrainTrainingSet <- trainingset[traintrainset, ]
TestTrainingSet <- trainingset[-traintrainset, ]
dim(TrainTrainingSet); dim(TestTrainingSet)
```
## The variable "classe" contains 5 levels: A, B, C, D and E. A plot of the outcome variable will allow us
## to see the frequency of each levels in the TrainTrainingSet data set and # compare one another
```{r}
plot(TrainTrainingSet$classe, col="blue", main="Plot of levels of variable classe within the Train", xlab="classe", ylab="Frequency")
```
# CrossValidation 1: Prediction with a Decision Tree

## Model
```{r}
model1 <- rpart(classe ~ ., data=TrainTrainingSet, method="class")
```

## Prediction
```{r}
prediction1 <- predict(model1, TestTrainingSet, type = "class")
```
## Plot the Decision Tree
```{r}
fancyRpartPlot(model1)
```

## Test results on our TestTrainingSet data set:
```{r}
cm <-confusionMatrix(prediction1, TestTrainingSet$classe)
print(cm)
```
## Accurate
```{r}
round(cm$overall['Accuracy'] * 100, 2)
```

## Error
```{r}
round(1 - cm$overall['Accuracy'],2)
```

# CrossValidation 2: Prediction with Random Forest

## Model
```{r}
model2 <- randomForest(classe ~. , data=TrainTrainingSet, method="class")
```

## Prediction
```{r}
prediction2 <- predict(model2, TestTrainingSet, type = "class")
```

## Test results on our TestTrainingSet data set:
```{r}
cm <-confusionMatrix(prediction2, TestTrainingSet$classe)
print(cm)
```

## Curve ROC
```{r}
plot(model2)
```

# CrossValidation 2: Prediction with Random Forest

## Decision on which Prediction Model to Use:

### Random Forest algorithm performed better than Decision Trees. Accuracy for Random Forest model was 0.995 (95% CI: (0.993, 0.997)) compared to Decision Tree model with 0.739 (95% CI: (0.727, 0.752)). The Random
### Forests model is choosen. The expected outofsample error is estimated at 0.005, or 0.5%.

# Submission
### Here is the final outcome based on the Prediction Model 2 (Random Forest) applied against the Testing dataset
```{r}
predictfinal <- predict(model2, testingset, type="class")
predictfinal
```
##Function to generate files with predictions to submit for assignment

```{r}
pml_write_files = function(x){
n = length(x)
for(i in 1:n){
filename = paste0("problem_id_",i,".txt")
write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
}
}
pml_write_files(predictfinal)
```
