---
title: "Practical Machine Learning - Prediction Assignment Writeup"
author: "David Hinostroza Vintes"
date: "17 de agosto de 2016"
output: html_document
---

For this assignment I analyzed the provided data to determine what activity an individual perform. To do this I made use of caret and randomForest, this allowed me to generate correct answers for each of the 20 test data cases provided in this assignment.

# Install packages and load the required libraries
```{r}
set.seed(1234)
library(lattice); 
library(ggplot2); 
library(caret); 
library(randomForest); 
library(rpart); 
library(rpart.plot);
library(rattle)
library(ROCR)
```

## The location where the training data is to be downloaded from
```{r}
trainUrl <-"http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
```
## The location where the testing data is to be downloaded from
```{r}
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
