---
title: "Practical Machine Learning - Prediction Assignment Writeup"
author: "David Hinostroza Vintes"
date: "17 de agosto de 2016"
output: html_document
---

For this assignment I analyzed the provided data to determine what activity an individual perform. To do this I made use of caret and randomForest, this allowed me to generate correct answers for each of the 20 test data cases provided in this assignment.

# Install packages and load the required libraries
```{r}
set.seed(1234)
library(lattice); 
library(ggplot2); 
library(caret); 
library(randomForest); 
library(rpart); 
library(rpart.plot);
library(rattle)
library(ROCR)
```

## The location where the training data is to be downloaded from
```{r}
trainUrl <-"pml-training.csv"
```
## The location where the testing data is to be downloaded from
```{r}
testUrl <- "pml-testing.csv"
```

## data load and clean up
```{r}
trainingset <- read.csv(trainUrl, na.strings=c("NA","#DIV/0!",""))
testingset <- read.csv(testUrl, na.strings=c("NA","#DIV/0!",""))
```

## Perform exploratory analysis
```{r}
dim(trainingset); 
dim(testingset); 
summary(trainingset); 
summary(testingset); 
str(trainingset); 
str(testingset); 
head(trainingset); 
head(testingset)
```

